---
title: "Forecast Model"
output: github_document
---

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE,
  fig.width = 14,
  fig.height = 6,
  out.width = "90%",
  set.seed(8))

library(tidyverse)
library(forecast)
library(astsa)
library(MLmetrics)
library(lubridate)
```

Time series forecasting involves using historical, time-stamped data to make predictions of what might happen in the future.

```{r original dataset}
# Load and clean dataset:

bakery_df = 
  read_csv("./Data/Bakery_sales.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    unit_price = str_replace(unit_price, "€", ""),
    unit_price = str_replace(unit_price, ",", "."),
    unit_price = as.numeric(unit_price),
    year = year(date),
    month = month(date),
    hour = hour(time),
    product_name = article) %>% 
  filter(product_name != ".") %>% 
  select(-article)
```

In 2021, how many products were sold each day?
```{r 2021 daily sales df}
sale_2021 = 
  bakery_df %>% 
  filter(year == 2021) %>% 
  select(date, year, month, hour, product_name, quantity, unit_price) %>% 
  group_by(date) %>% 
  summarize(
    total_sale = sum(quantity))

sale_2021
```

The `sale_2021` dataset shows the number of products sold each day in 2021. In 2021, the bakery opened for business for 339 days. 

The line plot below shows the total number of products sold each day in 2021.

```{r 2021 daily sales line plot}
sale_2021 %>% 
  ggplot(aes(x = date, y = total_sale)) +
  geom_line(aes(color = "Orange")) +
  scale_x_date(date_labels = "%b %Y", date_breaks  = "1 month") +
  labs(
    x = "Date",
    y = "Number of products sold",
    title = "Bakery's Daily Sales (2021)") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.4))
```

We will see this trend again as we verify the accuracy of different forecast models in the next few sections.


## Data Pre-processing

We need to create a simple and clean dataframe derived from the original dataset `bakery_df` that can be used by the forecast models. This new dataframe only needs two columns:                          
* `date` : date order in which the bakery opened for business                                                                                                                                               
* `total_sale` : sum of products sold 

The `sales_all` dataframe consists of 600 rows, in which the dates begin with *January 2, 2021* and end with *September 30, 2022*. 

```{r simplify df}
sales_all = 
  bakery_df %>% 
  group_by(date) %>% 
  summarize(
    total_sale = sum(quantity))
```

Then, we need to create a `train` and a `test` data frames so that we can test the accuracy of our forecast models. We will be using all data prior to *September 1, 2022* as our `train` dataset 
to predict the amount of products sold daily for September, 2022. The remaining data from *September 1, 2022* to *September 30, 2022* will be our `test` dataset.

```{r train test df}
train = 
  sales_all %>% 
  filter(date < "2022-09-01")

test = 
  sales_all %>% 
  filter(year(date) == 2022 & month(date) == 9)
```


### Mean Absolute Percentage Error (MAPE)

For model evaluation, we will be relying on the ***Mean Absolute Percentage Error (MAPE)*** to measure the accuracy of our predictions.

The accuracy of the model can be calculated as $100 - MAPE$. For example, an *MAPE* value of 70 indicates the model has an accuracy of 30%.


## Seasonal Naive Model

First, we begin with the most basic forecast model - the **Seasonal Naive Model**.

Naive forecasting is a simple and cost-effective method in which the forecasts produced are equal to the last observed value. The seasonal naive approach is used 
when the time series exhibits seasonality, in which case, the forecasts are equivalent to the value from the last season. Naive methods are typically used as a 
benchmark against which more sophisticated forecasting techniques can be compared. 

```{r SNM}
# Fit the model
seasonal_naive_model = snaive(train$total_sale, h = length(test$total_sale))

# Compute error of the model
SNM_error = MAPE(seasonal_naive_model$mean, test$total_sale) * 100
```

**The accuracy of the *Seasonal Naive* model is `r round(sum(SNM_error = 100 - SNM_error), digits = 2)`%.**

The *Seasonal Naive* model will output a list of mean values, and they are the predicted values of daily sales from September 1, 2022 to September 30, 2022.
We need to extract the list of mean values from the model and put them into the `test` dataframe so that we can produce a time-series forecast plot.

```{r SNM predict}
test_seasonal = 
  test %>% 
  mutate(
    pred_sale = seasonal_naive_model$mean)
```

Finally, we can plot a time-series forecasting graph to show the results of our *Seasonal Naive* model.

```{r plot SNM prediction}
train %>% 
  ggplot(aes(x = date, y = total_sale)) +
  geom_line(aes(color = "Actual sale (Prior to 2022-09-01)")) +
  geom_line(data = test_seasonal, aes(x = date, y = total_sale, color = "Actual sale (2022-09-01 to 2022-09-30)")) + 
  geom_line(data = test_seasonal, aes(x = date, y = pred_sale, color = "Predicted sale (2022-09-01 to 2022-09-30)"), size = 1.5) +
  scale_x_date(date_labels = "%b %Y", date_breaks  = "2 month") +
  labs(
    x = "Date",
    y = "Number of products sold",
    title = "Seasonal Naive Forecast of Daily Sales for September (2022)") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.4))
```

Comments: Although the *Seasonal Naive Model* gives high accuracy, the prediction line is a horizontal straight line, which means this model predicted that 
the amount of products sold is the same for each day. Therefore, we think that the predictions from this model are not the best.


## Double-Seasonal Holt-Winters (DSHW)

The **Double-seasonal Holt-Winters** method uses additive trend and multiplicative seasonality, where there are two seasonal components which are multiplied together. 
The length of the two seasonalities must be multiples of one another (2 and 4, 4 and 12, etc.). In our case, we will specify the length of the first seasonality as 7  
and the second as 14.

```{r DSHWM}
# Fit the model
double_seasonal_model = dshw(train$total_sale, period1 = 7, period2 = 14, h = length(test$total_sale))

# Compute the model's error
DSHW_error = MAPE(double_seasonal_model$mean, test$total_sale) * 100
```

**The accuracy of the *Double-seasonal Holt-Winters* model is `r round(sum(DSHW_error = 100 - DSHW_error), digits = 2)`%.**

The *DSHW* model will output a list of mean values, and they are the predicted values of daily sales from September 1, 2022 to September 30, 2022.
We need to extract the list of mean values from the model and put them into the `test` dataframe so that we can produce a time-series forecast plot.

```{r DSHW predict}
test_double_seasonal = 
  test %>% 
  mutate(
    pred_sale = double_seasonal_model$mean)
```

Finally, we can plot a time-series forecasting graph to see the results of our *DSHW* model.

```{r DSHW prediction plot}
train %>%
  filter(date > "2022-01-01") %>% 
  ggplot(aes(x = date, y = total_sale)) +
  geom_line(aes(color = "Actual sale (Prior to 2022-09-01)")) +
  geom_line(data = test_double_seasonal, aes(x = date, y = total_sale, color = "Actual sale (2022-09-01 to 2022-09-30)")) + 
  geom_line(data = test_double_seasonal, aes(x = date, y = pred_sale, color = "Predicted sale (2022-09-01 to 2022-09-30")) +
  scale_x_date(date_labels = "%b %Y", date_breaks  = "2 month") +
  labs(
    x = "Date",
    y = "Number of products sold",
    title = "Double-Seasonal Holt-Winters Forecast of Daily Sales for September (2022)") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.4))
```

Comments: Although the *DSHW* model has a worse accuracy than the basic model (*Seasonal Naive*), but visually it is better than the basic model because it shows dynamic trend.
This would make more sense than the predictions given by the *Seasonal Naive* model as we can see that it is no longer predicting that the daily sales are the same for each day.


## TBATS

**TBATS** is an acronym derived from the **Trigonometric seasonality, Box-Cox transformation, ARMA errors, Trend, and Seasonal components** of this approach. It takes its roots 
from exponential smoothing methods and is capable of modeling time series with multiple seasonalities.

```{r tbats model}
# Train a TBATS model
TBATS_model = tbats(train$total_sale)

# Generate forecast with the model
tbats_df = forecast(TBATS_model, h = length(test$total_sale))

# Check the error for the model
TBATS_error = MAPE(tbats_df$mean, test$total_sale) * 100
```

**The accuracy of the *TBATS* model is `r round(sum(TBATS_error = 100 - TBATS_error), digits = 2)`%.**

The *TBATS* model will output a list of mean values, and they are the predicted values of daily sales from September 1, 2022 to September 30, 2022.
We need to extract the list of mean values from the model and put them into the `test` dataframe so that we can produce a time-series forecast plot.

```{r TBATS predict}
test_tbats = 
  test %>% 
  mutate(
    pred_sale = tbats_df$mean)
```

```{r TBATS plot}
train %>% 
  filter(date > "2022-01-01") %>% 
  ggplot(aes(x = date, y = total_sale)) +
  geom_line(aes(color = "Actual sale (Prior to 2022-09-01)")) +
  geom_line(data = test_tbats, aes(x = date, y = total_sale, color = "Actual sale (2022-09-01 to 2022-09-30)")) + 
  geom_line(data = test_tbats, aes(x = date, y = pred_sale, color = "Predicted sale (2022-09-01 to 2022-09-30"), size = 0.8) +
  scale_x_date(date_labels = "%b %Y", date_breaks  = "2 month") +
  labs(
    x = "Date",
    y = "Number of products sold",
    title = "TBATS Forecast of Daily Sales for September (2022)") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.4))
```

Comments: As we can see from the plot, the prediction line (blue) demonstrates dynamic pattern as the *DSHW* model, but the accuracy of the *TBATS* model is slightly better than the *DSHW* model.


## Neural Network

A neural network is a series of algorithms that identifies patterns and relationships in data, similar to the way the brain operates. 

The `forecast` library comes with the option of a feed-forward neural network with a single hidden layer and lagged inputs for univariate time series forecasting.

```{r neural network model}
# Train a neural network model
nn_model = nnetar(train$total_sale)

# Generate forecast with the model
nn_forecast_df = forecast(nn_model, h = length(test$total_sale))

# Check the MAPE
NN_error = MAPE(nn_forecast_df$mean, test$total_sale) * 100
```

**The accuracy of the *Neural Network* model is `r round(sum(NN_error = 100 - NN_error), digits = 2)`%.**

The *Neural Network* model will output a list of mean values, and they are the predicted values of daily sales from September 1, 2022 to September 30, 2022.
We need to extract the list of mean values from the model and put them into the `test` dataframe so that we can produce a time-series forecast plot.

```{r NN prediction}
test_nn = 
  test %>% 
  mutate(
    pred_sale = nn_forecast_df$mean)
```

```{r NN plot}
train %>% 
# To show the prediction line more clearly, we are only showing the trends of daily sales after May 1, 2022.
  filter(date > "2022-05-01") %>% 
  ggplot(aes(x = date, y = total_sale)) +
  geom_line(aes(color = "Actual sale (Prior to 2022-09-01)")) +
  geom_line(data = test_nn, aes(x = date, y = total_sale, color = "Actual sale (2022-09-01 to 2022-09-30)"), alpha = 0.8) + 
  geom_line(data = test_nn, aes(x = date, y = pred_sale, color = "Predicted sale (2022-09-01 to 2022-09-30)"), size = 0.7, alpha = 0.8) +
  scale_x_date(date_labels = "%b %Y", date_breaks  = "1 month") +
  labs(
    x = "Date",
    y = "Number of products sold",
    title = "Neural Network Forecast of Daily Sales for September (2022)") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.4))
```

Comments: As shown in the plot, the predicted trendline (blue) looks similar to the actual trendline (red) highly overlaps it. Although the prediction accuracy is not 100%,
but this *Neural Network* model gives us the highest accuracy of prediction.


## Sales Forecast for Oct - Dec 2022

As we have identified that the *Neural Network* model gives us the highest accuracy (`r round(sum(NN_error = 100 - NN_error), digits = 2)`%), we will use this model to predict the future daily sales 
from October to December in 2022.

This time, we will use all the data from January 1, 2021 to September 30, 2022 to fit into the *Neural Network* forecasting model to predict the daily sales of October - December in 2022.

```{r future prediction}
# Train the model
nn_future = nnetar(sales_all$total_sale)

# Generate the forecast
nn_future_df = forecast(nn_future, h = 92)
```

Since predicted values given by the model do not comes with dates, we need to create a dataframe with specific range of date to combine with the predicted values.

```{r date df}
dates = seq(as.Date('2022-10-01'), as.Date('2022-12-31'), by = 'days')

date_df =
  tibble(
    A = letters[sample(1:26, 92, TRUE)])

date_df = 
  tibble(
    date = rep(dates, length.out = nrow(date_df)))

date_df =
  date_df %>% 
  mutate(
    pred_sale = nn_future_df$mean)
```

Then, we can plot a graph to show the predicted daily sales of the bakery from October to December in 2022. 

```{r predicted future sales}
date_df %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = pred_sale, color = "Red")) +
  scale_x_date(date_labels = "%b %d", date_breaks  = "10 day") +
  labs(
    x = "Date (Month Day)",
    y = "Number of products sold each day",
    title = "Predicted Daily Sales from October to December (2022)") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.4))
```
