---
title: "Forecast Model"
output: github_document
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)
library(forecast)
library(astsa)
library(MLmetrics)
library(zoo)
library(lubridate)
```

Time series forecasting involves using historical, time-stamped data to make predictions of what might happen in the future.

Load and clean dataset:
```{r}
bakery_df = 
  read_csv("./Data/Bakery_sales.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    unit_price = str_replace(unit_price, "€", ""),
    unit_price = str_replace(unit_price, ",", "."),
    unit_price = as.numeric(unit_price),
    year = year(date),
    month = month(date),
    hour = hour(time),
    product_name = article) %>% 
  filter(product_name != ".") %>% 
  select(-article)
```

In 2021, how many products were sold each day?
```{r}
sale_2021 = 
  bakery_df %>% 
  filter(year == 2021) %>% 
  select(date, year, month, hour, product_name, quantity, unit_price) %>% 
  group_by(date) %>% 
  summarize(
    total_sale = sum(quantity))
```

In 2021, the bakery opened for business for 339 days. The `sale_2021` dataset shows the number of products sold each day in 2021. 
To begin our forecasting task, we need to convert the dataframe into a *time series* or *ts* object. 

Plot data on a line graph
```{r}
sale_2021 %>% 
  ggplot(aes(x = date, y = total_sale)) +
  geom_line()
```


## Seasonal Naive Model = Baseline Forecast

Naive forecasting is a simple and cost-effective method in which the forecasts produced are equal to the last observed value. The seasonal naive approach is used 
when the time series exhibits seasonality, in which case, the forecasts are equivalent to the value from the last season. Naive methods are typically used as a 
benchmark against which more sophisticated forecasting techniques can be compared. 
For model evaluation, we will be relying on the Mean Absolute Percentage Error (MAPE) to measure the accuracy of our predictions.

```{r}
sales_all = 
  bakery_df %>% 
  group_by(date) %>% 
  summarize(
    total_sale = sum(quantity))
```

```{r}
train = 
  sales_all %>% 
  filter(year(date) == 2021)

test = 
  sales_all %>% 
  filter(year(date) == 2022 & month(date) == 1)
```

```{r}
seasonal_naive_model = snaive(train$total_sale, h = length(test$total_sale))

MAPE(seasonal_naive_model$mean, test$total_sale) * 100
```

```{r}
test_seasonal = 
  test %>% 
  mutate(
    pred_sale = seasonal_naive_model$mean)
```

```{r}
train %>% 
  ggplot(aes(x = date, y = total_sale)) +
  geom_line(aes(color = "Actual sale (2021)")) +
  geom_line(data = test_seasonal, aes(x = date, y = total_sale, color = "Actual sale (2022)")) + 
  geom_line(data = test_seasonal, aes(x = date, y = pred_sale, color = "Predicted sale"), size = 1.5) +
  labs(
    x = "Date",
    y = "Number of products sold",
    title = "Seasonal Naive Forecast for January, 2022")
```

This is the most shit forecast model...


## Double-Seasonal Holt-Winters (DSHW)

```{r}
double_seasonal_model = dshw(train$total_sale, period1 = 7, period2 = 14, h = length(test$total_sale))

print("The error rate of this model is")

MAPE(double_seasonal_model$mean, test$total_sale) * 100
```

```{r}
test_double_seasonal = 
  test %>% 
  mutate(
    pred_sale = double_seasonal_model$mean)
```

```{r}
train %>% 
  ggplot(aes(x = date, y = total_sale)) +
  geom_line(aes(color = "Actual sale (2021)")) +
  geom_line(data = test_double_seasonal, aes(x = date, y = total_sale, color = "Actual sale (2022)")) + 
  geom_line(data = test_double_seasonal, aes(x = date, y = pred_sale, color = "Predicted sale"), size = 0.8) +
  labs(
    x = "Date",
    y = "Number of products sold",
    title = "Double-Seasonal Holt-Winters Forecast for January, 2022")
```

This model is better than the basic one, but still shit...


## TBATS

```{r}
TBATS_model = tbats(train$total_sale)

tbats_df = forecast(TBATS_model, h = length(test$total_sale))

print("The error rate of this model is")

MAPE(tbats_df$mean, test$total_sale) * 100
```

```{r}
test_tbats = 
  test %>% 
  mutate(
    pred_sale = tbats_df$mean)
```

```{r}
train %>% 
  ggplot(aes(x = date, y = total_sale)) +
  geom_line(aes(color = "Actual sale (2021)")) +
  geom_line(data = test_tbats, aes(x = date, y = total_sale, color = "Actual sale (2022)")) + 
  geom_line(data = test_tbats, aes(x = date, y = pred_sale, color = "Predicted sale"), size = 0.8) +
  labs(
    x = "Date",
    y = "Number of products sold",
    title = "TBATS Forecast for January, 2022")
```

