---
title: "Forecast Model"
output: github_document
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)
library(forecast)
library(astsa)
library(MLmetrics)
library(lubridate)
```

Time series forecasting involves using historical, time-stamped data to make predictions of what might happen in the future.

Load and clean dataset:
```{r}
bakery_df = 
  read_csv("./Data/Bakery_sales.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    unit_price = str_replace(unit_price, "€", ""),
    unit_price = str_replace(unit_price, ",", "."),
    unit_price = as.numeric(unit_price),
    year = year(date),
    month = month(date),
    hour = hour(time),
    product_name = article) %>% 
  filter(product_name != ".") %>% 
  select(-article)
```

In 2021, how many products were sold each month?
```{r}
sale_2021 = 
  bakery_df %>% 
  filter(year == 2021) %>% 
  select(date, year, month, hour, product_name, quantity, unit_price) %>% 
  group_by(date) %>% 
  summarize(
    total_sale = sum(quantity))
```

In 2021, the bakery opened for business for 339 days. The `sale_2021` dataset shows the number of products sold each day in 2021. 
To begin our forecasting task, we need to convert the dataframe into a *time series* or *ts* object. 

```{r}
ts_df = 
  ts(sale_2021$total_sale, start = c(2021), frequency = 339)

print(ts_df)
```

Plot data on a line graph
```{r}
plot.ts(ts_df, col = "red", main = "Monthly Sale (2021)", xlab = "Month", ylab = "Number of prroducts sold")
```



